Lapse & Vision Loss PS Analysis

### 2025-08-06 analysis.Rmd
wip: reviewer response

source("settings.R")

library(broom)
library(cobalt)
library(dplyr)
library(flextable)
library(gee)
library(ggplot2)
library(gt)
library(gtsummary)
library(kableExtra)
library(jtools)
library(margins)
library(MatchIt)
library(purrr)
library(readr)
library(survey)
library(twang)
library(WeightIt)

@@

data <- readr::read_csv(file.path(s_root, "Gina's Project/df_Final_240909.csv")) |> 
  mutate(
    outcome_va_vi = outcome_VA_logMAR >= 0.3,
    outcome_va_blind = outcome_VA_logMAR >= 1,
    outcome_va_vi_binary = if_else(outcome_va_vi, 1, 0),
    outcome_va_blind_binary = if_else(outcome_va_blind, 1, 0),
    ever_lapse_cat = factor(person_ever_lapse),
    ever_lapse_binary = if_else(ever_lapse_cat == TRUE, 1, 0),
    MRN = dplyr::min_rank(e_mrn_deidentified),
    gender_cat = factor(gender),
    race_ethnic_cat = factor(race_ethnicity_gp),
    insurance_cat = factor(insurance_gp),
    age_cat = cut(age, breaks = c(0, 20, 45, 65, Inf), labels = c("0-20", "21-45", "46-65", "65+")),
    person_dr = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 1,
      NPDR == "Severe" ~ 1,
      PDR == "Present" ~ 2,
    ),
    person_dr_severity = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 2,
      NPDR == "Severe" ~ 3,
      PDR == "Present" ~ 4,
    ),
    person_ever_treat = case_when(
      other_inject == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & PRP_flag == 0 ~ FALSE,
      other_inject == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | PRP_flag == 1 ~ TRUE
    ),
    glaucoma_bef_hitplus_cat = case_when(
      Glaucoma_bef_hitplus != "Present" | is.na(Glaucoma_bef_hitplus) ~ FALSE,
      Glaucoma_bef_hitplus == "Present" ~ TRUE
    ),
    glaucoma_after_hitplus_cat = case_when(
      Glaucoma_after_hitplus != "Present" | is.na(Glaucoma_after_hitplus) ~ FALSE,
      Glaucoma_after_hitplus == "Present" ~ TRUE
    ),
    otherretina_bef_hitplus_cat = case_when(
      Otherretina_bef_hitplus != "Present" | is.na(Otherretina_bef_hitplus) ~ FALSE,
      Otherretina_bef_hitplus == "Present" ~ TRUE
    ),
    otherretina_after_hitplus_cat = case_when(
      Otherretina_after_hitplus != "Present" | is.na(Otherretina_after_hitplus) ~ FALSE,
      Otherretina_after_hitplus == "Present" ~ TRUE
    ),
    other_inject = factor(other_inject),
    anti_VEGF = factor(anti_VEGF),
    focal_laser_flag = factor(focal_laser_flag),
    PRP_flag = factor(PRP_flag),
    glaucoma_bef_hitplus_cat = factor(glaucoma_bef_hitplus_cat),
    otherretina_bef_hitplus_cat = factor(otherretina_bef_hitplus_cat),
    catsurg_before_hitplus_cat = factor(cataract_surgery_bef_hitplus)
  )

save_og_image <- function(plot, filename, width=800, height=1000) {
  png(filename=filename, width=width, height=height)
  plot
  dev.off()
}

@@

```{r get_m_data}
m_data <- data |> filter(
  !is.na(CCI) & !is.na(DCSI)
)

@@

fit_logistic_models <- function(design, formulas) {
  models <- list()
  for (i in seq_along(formulas)) {
    model <- svyglm(
      formula = formulas[[i]],
      design = design,
      family = quasibinomial(),
      contrasts = NULL
    )
    models[[names(formulas)[i]]] <- model
  }
  models
}

summarize_logistic_models <- function(models, method_name) {
  roundTo <- 2
  roundP <- 3

  results <- lapply(names(models), function(name) {
    model <- models[[name]]

    model_summary <- summary(model)
    coef_table <- model_summary$coefficients
    estimate <- coef_table[2, "Estimate"]

    or <- formatC(exp(estimate), format = "f", digits = roundTo)
    estimate <- formatC(estimate, format = "f", digits = roundTo)
    se <- formatC(coef_table[2, "Std. Error"], format = "f", digits = roundTo)
    t_value <- formatC(coef_table[2, "t value"], format = "f", digits = roundTo)
    p_value <- formatC(coef_table[2, "Pr(>|t|)"], format = "f", digits = roundP)

    confint_values <- confint(model, level = 0.95)[2, ]
    ci_lb_or <- formatC(exp(confint_values[1]), format = "f", digits = roundTo)
    ci_ub_or <- formatC(exp(confint_values[2]), format = "f", digits = roundTo)

    package <- if (method_name %in% c("cbps", "entropy", "bart")) "WeightIt"
    else if (method_name %in% c("twang_gbm")) "twang"
    else "MatchIt"

    approach <- if (method_name %in% c("cbps", "entropy", "bart", "twang_gbm")) "ipw"
    else if (method_name == "subclass_glm") "subclassification"
    else "match w/ replacement"

    tibble::tibble(
      package = package,
      method = method_name,
      approach = approach,
      estimand = "ATT",
      formula = name,
      estimate = estimate,
      estimate_or = or,
      se = se,
      t_value = t_value,
      p_value = p_value,
      lb_95_or = ci_lb_or,
      ub_95_or = ci_ub_or
    )
  })

  bind_rows(results)
}

@@

tidy_output_model_formatted <- function(model, digits = 2) {
  tidy(model, exponentiate = TRUE, conf.int = TRUE) |>
    mutate(
      estimate = round(estimate, digits),
      std.error = round(std.error, digits),
      statistic = round(statistic, digits),
      conf.low = round(conf.low, digits),
      conf.high = round(conf.high, digits),
      p.value = round(p.value, digits + 1),
      # OR_CI = paste0(estimate, " (", conf.low, ", ", conf.high, ")"),
      p.value.sum = case_when(
        p.value < 0.001 ~ "<.001",
        TRUE ~ formatC(p.value, format="f", digits = 3)
      )
    )
}
tidy_output_model <- tidy_output_model_formatted(output_model)
tidy_output_model
flextable(tidy_output_model) |> save_as_docx(path = file.path(ps_dir, paste0("output_model_tidy_formatted.docx")))

>>>>

source("settings.R")
source("functions.R")

# Load all required libraries
load_required_libraries()

@@

# Load and prepare data
data_path <- file.path(s_root, "Gina's Project/df_Final_240909.csv")
data <- load_and_prepare_data(data_path)

# save_og_image function is now loaded from functions.R

@@

```{r get_m_data}
# Apply exclusions (removes missing CCI/DCSI)
m_data <- apply_exclusions(data)

@@

# fit_logistic_models function is now loaded from functions.R

# summarize_logistic_models function is now loaded from functions.R

@@

# tidy_output_model_formatted function is now loaded from functions.R
tidy_output_model <- tidy_output_model_formatted(output_model)
tidy_output_model
flextable(tidy_output_model) |> save_as_docx(path = file.path(ps_dir, paste0("output_model_tidy_formatted.docx")))
